---
title: "Pharmaceutical machine learning with tidymodels"
description: |
  Use Machine Learning to develop a model to determine if a proposed drug could be a mutage.
date: 2024-03-03
categories:
  - classification
  - logistic regression
  - random forest
  - xgboost
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    toc_float: true    
    self_contained: false
editor_options: 
  chunk_output_type: console
draft: True
---
```{css, echo = FALSE}
d-byline {
  display: none;
  margin:0;
}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(results = FALSE, echo = TRUE, eval = FALSE, fig.width = 6, layout="l-body-outset")
```

### Introduction

I learnt about this dataset while watching [Ep. 7: End-to-end machine learning workflow with tidymodels & Posit Team](https://youtu.be/O0Dklq-IZhw?si=q_Gph6nE6jOBOx4a) on Posit's youtube channel. 

This episode was hosted by [Simon Couch](https://www.simonpcouch.com/) whom I first learnt about while researching about tidymodel `stacks`. Here is a link to Simon's [github](https://github.com/simonpcouch/mutagen) page where you can obtain the dataset for practice.

Here is a brief description of the project. A group of scientists investigate whether they can use drug information to predict if a proposed drug could be a mutagen (i.e., toxicity caused by damage to DNA). In pharmaceutical research, mutagenicity refers to a drug’s tendency to increase the rate of mutations due to the damage of genetic material, a key indicator that a drug may be a carcinogen. Mutagenicity can be evaluated using a lab test, though the test requires experienced scientists and time in the lab. A group of scientists are studying whether, instead, they can use known information to quickly predict the mutagenicity of new drugs.

### Importing the data

```{r, eval = TRUE}
rm(list=ls())
pacman::p_load("tidyverse", #for tidy data science practice
               "tidymodels", "workflows", "finetune", "themis", "embed", "butcher",# for tidy machine learning
               "pacman", #package manager
               "devtools", #developer tools
               "Hmisc", "skimr", "broom", "modelr",#for EDA
               "jtools", "huxtable", "interactions", # for EDA
               "ggthemes", "ggstatsplot", "GGally",
               "scales", "gridExtra", "patchwork", "ggalt", "vip",
               "ggstance", "ggfortify", # for ggplot
               "DT", "plotly", #interactive Data Viz
               # Lets install some ML related packages that will help tidymodels::
               "usemodels", "poissonreg", "agua", "sparklyr", "dials",#load computational engines
               "doParallel", # for parallel processing (speedy computation)
               "ranger", "xgboost", "glmnet", "kknn", "earth", "klaR", "discrim", "naivebayes", "baguette", "kernlab",#random forest
               "janitor", "lubridate")
load("mutagen_results.RData")
```

```{r}
df <-read_csv("mutagen.csv")
df %>% count(outcome)
data <-
  df %>% 
  dplyr::select(-1)
# confirm no NA data
table(is.na(data))
```

There are no missing fields. This is a "wide" data set, with 1581 columns. There is no missing data, so we do not have to address it in a recipe step. I will skip the EDA portion, and move straight into splitting the data.

### Splitting the Data

I split the data into a training and a test set, strata by outcome. I also create a cross validation dataset for tuning hyper-parameters.

```{r}

set.seed(2024030101)
data_split <-
  data %>% 
  initial_split(strata = outcome)

data_train <-
  data_split %>% 
  training()
data_test <-
  data_split %>% 
  testing()

data_fold <-
  data_train %>% 
  vfold_cv(v = 10, strata = outcome)

```

### Create recipes

I will evaluate 2 recipes, a base recipe and another with PCA.

```{r}
base_rec <-
  recipes::recipe(formula = outcome ~.,
                  data = data_train) %>% 
  step_zv(all_predictors()) %>% # remove zero variance
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors())

pca_rec <-
  base_rec %>% 
  step_pca(all_predictors())
```

### Specify models

I will evaluate the performance of 3 models, random forest(I'm not sure why Simon didn't make use of a random forest model), xgboost, and logistic regression.

```{r}

# random forest
rf_spec <-
  rand_forest() %>% 
  set_engine("ranger",
             importance = "impurity") %>% 
  set_mode("classification") %>% 
  set_args(trees = tune(),
           mtry = tune(),
           min_n = tune())

# xgboost
xgb_spec <-
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") %>% 
  set_args(trees = 1000L,
           tree_depth = tune(),
           min_n = tune(),
           loss_reduction = tune(),
           sample_size = tune(),
           mtry = tune(),
           learn_rate = tune(),
           stop_iter = 10)

# Logistic Regression Model
logistic_spec <- 
  logistic_reg() %>%
  set_engine(engine = 'glm') %>%
  set_mode('classification') 

null_spec <-
  null_model() %>% 
  set_mode("classification") %>% 
  set_engine("parsnip")

```

### Workflowset

I created a workflowset, which will conveniently allow me to evaulate the models across both recipes.

```{r}

# workflow set

base_set <- 
  workflow_set (
    list(basic = base_rec,
         pca = pca_rec), #preprocessor
    list(rand_forest = rf_spec,
         xgboost = xgb_spec,
         logistic = logistic_spec,
         null = null_spec), #model
    cross = TRUE) #default is cross = TRUE

```

### Tune hyper-parameters using `tune_race_anova`

Let's tune hyper-parameters using `tune_race_anova`, setting `registerDoParallel` for speedy processing.

```{r}
set.seed(2024030302)
doParallel::registerDoParallel(cl=15, cores = 30)

racing_results <-
  workflow_map(base_set,
               fn = "tune_race_anova",
               resamples = data_fold,
               metrics = metric_set(roc_auc, f_meas, accuracy, mn_log_loss),
               control = control_race(verbose = TRUE,
                                      verbose_elim = TRUE,
                                      allow_par = TRUE,
                                      save_workflow = FALSE,
                                      parallel_over = "everything"))

```

Once racing is done, we can use `autoplot` to see which was the "winning" model.

```{r, eval = TRUE, preview = TRUE}
autoplot(racing_results) + theme_bw() + theme(legend.position = "bottom")
```

Let's rank the results by "roc_auc".

```{r, eval = TRUE}
racing_results %>% 
  workflowsets::rank_results(rank_metric = "roc_auc") %>% 
  filter(.metric == "roc_auc") %>% 
  dplyr::select(wflow_id, mean, std_err, rank) %>% 
  datatable() %>% 
  formatRound(columns = c("mean", "std_err"),
              digits = 3)
```
We see that the best performing model is a random forest model with a base recipe. It has a roc_auc score of 0.901. Let's take a look at its hyper-parameters.

```{r}
rf_tuned_parameters <-
  racing_results %>% 
  extract_workflow_set_result(id = "basic_rand_forest") %>% 
  select_best(metric = "roc_auc")
```

Hyper-parameters are mtry of 1146 and min_n of 14. Let's see if we can improve on these results by performing iterative search using `tune_sim_anneal`.

```{r}
# base recipe and rf workflow
base_rf_wflow <-
  workflow() %>% 
  add_model(rf_spec) %>% 
  add_recipe(base_rec)

# set up new search grid
rf_grid <-
  extract_parameter_set_dials(rf_spec) %>% 
  update(trees = trees(c(500L,2000L)),
         mtry = mtry(c(1000L,1500L)),
         min_n = min_n(c(5L,20L)
                       )
         )

rf_metrics <-
  metric_set(roc_auc, f_meas, accuracy, mn_log_loss)
```

Let's search!

### Iterative search using `tune_sim_anneal`

```{r}
# tune_sim_anneal

set.seed(2024030303)
doParallel::registerDoParallel(cl=15, cores = 30)

base_rf_sim_anneal_result <-
  tune_sim_anneal(
    object = base_rf_wflow,
    resamples = data_fold,
    iter = 25,
    metrics = rf_metrics,
    param_info = rf_grid,
    initial = 1,
    control = control_sim_anneal(verbose = TRUE,
                                 verbose_iter = TRUE,
                                 allow_par = TRUE,
                                 parallel_over = "everything")
  )
```

The search is done, let's take a look at the results

```{r eval = TRUE}
base_rf_sim_anneal_result %>% 
  collect_metrics(summarize = FALSE) %>% 
  filter(.metric =="roc_auc") %>% 
  arrange(desc(.estimate)) %>% 
  datatable()
```

Yes! I have been able to improve the model performance, achieving an `roc_auc` of 0.917 with hyper-parameters: 791 trees, mtry 1329 and min_n 9. Let's finalize the workflow


