---
title: "Pharmaceutical machine learning with tidymodels"
description: |
  Use Machine Learning to develop a model to determine if a proposed drug could be a mutagen.
date: 2024-03-03
categories:
  - classification
  - random forest
  - tune_grid
  - tune_bayes
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    toc_float: true    
    self_contained: false
editor_options: 
  chunk_output_type: console
draft: True
---
```{css, echo = FALSE}
d-byline {
  display: none;
  margin:0;
}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(results = FALSE, echo = TRUE, eval = FALSE, fig.width = 6, layout="l-body-outset")
```

### Introduction

I learnt about this dataset while watching [Ep. 7: End-to-end machine learning workflow with tidymodels & Posit Team](https://youtu.be/O0Dklq-IZhw?si=q_Gph6nE6jOBOx4a) on Posit's youtube channel. You can also read about it on Posit's [blogpost](https://posit.co/blog/pharmaceutical-machine-learning-with-tidymodels-and-posit-connect/).

This episode was hosted by [Simon Couch](https://www.simonpcouch.com/) whom I first learnt about while researching about tidymodel `stacks`. Here is a link to Simon's [github](https://github.com/simonpcouch/mutagen) page where you can obtain the dataset for practice.

Here is a brief description of the project. A group of scientists investigate whether they can use drug information to predict if a proposed drug could be a mutagen (i.e., toxicity caused by damage to DNA). In pharmaceutical research, mutagenicity refers to a drug’s tendency to increase the rate of mutations due to the damage of genetic material, a key indicator that a drug may be a carcinogen. Mutagenicity can be evaluated using a lab test, though the test requires experienced scientists and time in the lab. A group of scientists are studying whether, instead, they can use known information to quickly predict the mutagenicity of new drugs.

### Importing the data

```{r, eval = TRUE}
rm(list=ls())
pacman::p_load("tidyverse", #for tidy data science practice
               "tidymodels", "workflows", "finetune", "themis", "embed", "fastICA", "dimRed", "uwot", "butcher", # for tidy machine learning
               "pacman", #package manager
               "devtools", #developer tools
               "Hmisc", "skimr", "broom", "modelr",#for EDA
               "jtools", "huxtable", "interactions", # for EDA
               "ggthemes", "ggstatsplot", "GGally",
               "scales", "gridExtra", "patchwork", "ggalt", "vip",
               "ggstance", "ggfortify", # for ggplot
               "DT", "plotly", #interactive Data Viz
               # Lets install some ML related packages that will help tidymodels::
               "usemodels", "poissonreg", "agua", "sparklyr", "dials",#load computational engines
               "doParallel", # for parallel processing (speedy computation)
               "ranger", "xgboost", "glmnet", "kknn", "earth", "klaR", "discrim", "naivebayes", "baguette", "kernlab",#random forest
               "janitor", "lubridate")
load("mutagen_results.RData")

```

```{r}
df <-read_csv("mutagen.csv")
df %>% count(outcome)
data <-
  df %>% 
  dplyr::select(-1) %>% 
  mutate(outcome = fct_relevel(outcome, "mutagen")) %>% 
  janitor::clean_names()
# confirm no NA data
table(is.na(data))
```

This is a "wide" data set, with `r ncol(data)` columns and `r nrow(data)` rows. It appears to be an ideal dataset to investigate dimensionality reduction and feature extraction, which I will be learning about next week in Module 3 of my class.

There is no missing data, so we do not have to address it in a recipe step. I will skip the EDA portion, and move straight into splitting the data.

A note for those who intend to follow along, or make use of this dataset. It is computationally intensive. My desktop running a ryzen 3950X with 64gb ram ran `first_tune` in approx 4 hours. Subsequently, 10 iterations of `tune_bayes` took approx 12 hours.

### Splitting the Data

I split the data into a training and a test set, strata by outcome. I also create a cross validation dataset for tuning hyper-parameters.

```{r}

set.seed(2024030101)
data_split <-
  data %>% 
  initial_split(strata = outcome)

data_train <-
  data_split %>% 
  training()
data_test <-
  data_split %>% 
  testing()
data_fold <-
  data_train %>% 
  vfold_cv(v = 10, strata = outcome)

```

### Modeling approach

Simon approached this problem by evaluating several models, then fine-tuning the chosen model, xgboost, using `tune_sim_anneal`. Random Forest wasn't among the models he evaluated. 
I will approach this problem slightly differently. I will evaluate 4 recipes: the first is a "basic" `rec_base` with all features. Then, I will create 3 other recipes which use different methods for dimension reduction and feature extraction. They are: 
-   `rec_pca` where I will use `step_pca` for principal component analysis method,
-   `rec_ica` where I will use `step_ica` for independent component analysis method,
-   `rec_umap` where I will use `step_umap` for uniform manifold approximation and projection method.

I also also evaluate only one model, the random forest model, and tune its hyper-parameters using `tune_grid`.

Finally, I will fine-tune the hyper-parameters further using `tune_bayes`.

### Create recipes

Let's create the 4 recipes.

```{r}
rec_base <-
  recipes::recipe(formula = outcome ~.,
                  data = data_train) %>% 
  step_zv(all_predictors()) %>% # remove zero variance
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors())

rec_pca <-
  rec_base %>% 
  step_pca(all_numeric_predictors(), num_comp = tune(), threshold = tune())
# 
# rec_pls <-
#   rec_base %>% 
#   step_pls(all_numeric_predictors(), outcome = "outcome", num_comp = tune())

rec_ica <-
  rec_base %>% 
  step_ica(all_numeric_predictors(), num_comp = tune())

rec_umap <-
  rec_base %>% 
  step_umap(all_numeric_predictors(),
            outcome = "outcome",
            num_comp = tune(),
            neighbors = tune(),
            min_dist = tune(),
            learn_rate = tune(),
            epochs = tune())

```

### Specify models

I will evaluate the performance of the random forest model.

```{r}

# random forest
spec_rf <-
  rand_forest() %>% 
  set_engine("ranger",
             importance = "impurity") %>% 
  set_mode("classification") %>% 
  set_args(trees = tune(),
           mtry = tune(),
           min_n = tune())

```

### Workflowset

I created a workflowset, which will conveniently allow me to evaluate the random forest model across both 4 recipes. It was necessary to set parameter ranges for the various hyper-parameters that were going to be tuned. I ran into errors without this step.

```{r}

# workflow set
base_set <- 
  workflow_set (
    list(ica = rec_ica,
         pca = rec_pca,
         #PLS = rec_pls,
         umap = rec_umap,
         basic = rec_base), #preprocessor
    list(rf = spec_rf), #model
    cross = TRUE) #default is cross = TRUE

# set parameters

param_basic <-
  base_set %>%
  extract_workflow(id = "basic_rf") %>% 
  extract_parameter_set_dials() %>% 
  finalize(data_train)

param_ica <-
  base_set %>%
  extract_workflow(id = "ica_rf") %>% 
  extract_parameter_set_dials() %>% 
  finalize(data_train)

param_pca <-
  base_set %>%
  extract_workflow(id = "pca_rf") %>% 
  extract_parameter_set_dials() %>% 
  finalize(data_train)

param_umap <-
  base_set %>%
  extract_workflow(id = "umap_rf") %>% 
  extract_parameter_set_dials() %>% 
  finalize(data_train)

base_set <-
  base_set %>% 
  option_add(param_info = param_basic, id = "basic_rf") %>% 
  option_add(param_info = param_ica, id = "ica_rf") %>% 
  option_add(param_info = param_pca, id = "pca_rf") %>% 
  option_add(param_info = param_umap, id = "umap_rf")


```

### Tune hyper-parameters using `tune_grid`

Let's tune hyper-parameters using `tune_grid`, setting `registerDoParallel` for speedy processing.

```{r}
set.seed(2024030302)
cl <- (detectCores()/2) - 1
cores <- cl*2

doParallel::registerDoParallel(cl, cores)

first_tune <-
  workflow_map(base_set,
               fn = "tune_grid",
               verbose = TRUE,
               seed = 2024030302,
               grid = 11,
               resamples = data_fold,
               metrics = metric_set(roc_auc, accuracy),
               control = control_grid(verbose = TRUE,
                                      allow_par = TRUE,
                                      parallel_over = "everything"))
save(first_tune, file = "first_tune.Rda")
```

Once tuning is done, we can use `autoplot` to view the results.

```{r, eval = TRUE, preview = TRUE}
autoplot(first_tune) + theme_bw() + theme(legend.position = "bottom")
```

Let's rank the results by "roc_auc".

```{r, eval = TRUE, results = TRUE}
first_tune %>% 
  workflowsets::rank_results(rank_metric = "roc_auc") %>% 
  filter(.metric == "roc_auc") %>% 
  dplyr::select(wflow_id, mean, std_err, rank) %>% 
  datatable() %>% 
  formatRound(columns = c("mean", "std_err"),
              digits = 3)
```

We see that the best performing model is a random forest model with a base recipe. It has a roc_auc score of 0.900. Let's take a look at its hyper-parameters. I'm surprised that none of the other recipes made in into the top 10 positions. The next best recipe rf combination was `pca_rf` with a roc_auc score of 0.87.

```{r}
first_tune_param <-
  first_tune %>% 
  extract_workflow_set_result(id = "pca_rf") %>% 
  select_best(metric = "roc_auc")
```

Hyper-parameters are mtry of 1546 and min_n of 31. Let's see if we can improve on these results by performing iterative search using `tune_bayes`.

```{r}
# base recipe and rf workflow
basic_rf_res <-
  first_tune %>% 
  extract_workflow_set_result(id = "basic_rf")
basic_rf_wflow <-
  workflow() %>% 
  add_recipe(rec_base) %>% 
  add_model(spec_rf)

set.seed(2024030303)
cl <- (detectCores()/2) - 1
cores <- cl*2

doParallel::registerDoParallel(cl, cores)

bayes_tune <-
  basic_rf_wflow %>% 
  tune_bayes(
  resamples = data_fold,
  metrics = metric_set(roc_auc, accuracy),
  initial = basic_rf_res,
  iter = 25,
  param_info = param_basic,
  control = control_bayes(verbose = TRUE,
                          verbose_iter = TRUE,
                          no_improve = 10L,
                          allow_par = TRUE,
                          parallel_over = "everything")
)

save(bayes_tune, file = "bayes_tune.Rda")
```

The search is "done". Well, I actually ended it early as I got tired of waiting. My PC had already been on for over 36 hours and electricity is expensive! Let's take a look at the results.

```{r, eval = TRUE, results = TRUE}
bayes_tune %>% 
  collect_metrics(summarize = FALSE) %>% 
  dplyr::select(-.estimator) %>% 
  filter(.metric =="roc_auc") %>% 
  arrange(desc(.estimate)) %>% 
  datatable() %>% 
  formatRound(columns = ".estimate",
              digits = 3)
```

Yes! I have been able to improve the model performance, achieving an `roc_auc` of 0.918 with hyper-parameters: 1962 trees, mtry 826 and min_n 2. Let's finalize the workflow

### Last_fit and collect predictions

Let's extract the workflow and `last_fit` to `data_split`.

```{r}
base_rf_tuned_workflow <-
  basic_rf_wflow %>% 
  finalize_workflow(bayes_tune %>% select_best(metric = "roc_auc"))

base_rf_fit <-
  base_rf_tuned_workflow %>% 
  last_fit(data_split)
```

Let's compare it's metrics on the test set vs training set.

```{r eval = TRUE, results = TRUE}
base_rf_fit %>% collect_metrics()
```

We can also collect predictions and plot a roc_auc curve.

```{r}
pred_rf <-
  base_rf_fit %>% 
  collect_predictions() %>% 
  mutate(algo = "Basic RF")

```

```{r, results = TRUE, eval = TRUE}
pred_rf %>% 
  roc_curve(outcome, .pred_mutagen) %>% 
  autoplot()
```

### Which features were important?

Let's view the features by importance using `extract_fit_parsnip` and `vip`

```{r, results = TRUE, eval = TRUE}
base_rf_fit %>% 
  extract_fit_parsnip() %>% 
  vip(geom = "col",
      num_features = 20L) +
  labs(x = "Importance",
       y = "Features",
       title = "Features of Importance - Basic Recipe Random Forest Model") +
  theme_bw()
```

### Material added 10 March 2024

In a piece of homework completed recently [Using ML to predict customer attrition](https://yeokcmark.netlify.app/ml/2024-03-07-using-ml-to-predict-customer-attrition/), a simple Logistic Regression model was able to outperform an xgboost and K-nearest neighbor model when tested against training data.

I wonder how a Logistic Regression model would perform here? Let's investigate.

```{r}
# Logistic Regression Model
logistic_spec <- 
  logistic_reg() %>%
  set_engine(engine = 'glm') %>%
  set_mode('classification') 

logistic_set <-
    workflow_set (
    list(ica = rec_ica,
         pca = rec_pca,
         #PLS = rec_pls,
         umap = rec_umap,
         basic = rec_base), #preprocessor
    list(logistic = logistic_spec
    ), #model
    cross = TRUE) #default is cross = TRUE

set.seed(2024030304)
cl <- (detectCores()/2) - 1
cores <- cl*2
doParallel::registerDoParallel(cl, cores)

logistic_tune <-
  workflow_map(logistic_set,
               fn = "tune_grid",
               verbose = TRUE,
               seed = 2024030302,
               grid = 11,
               resamples = data_fold,
               metrics = metric_set(roc_auc, accuracy),
               control = control_grid(verbose = TRUE,
                                      allow_par = TRUE,
                                      parallel_over = "everything"))
save(logistic_tune, file = "logistic_tune.Rda")


```

The performance on training data isnt too shabby, and it actually favored the `rec_pca` recipe. It achieved a roc_auc of 0.912.

```{r, eval = TRUE, results = TRUE}
logistic_tune %>% 
  collect_metrics(summarize = FALSE) %>% 
  dplyr::select(-.estimator) %>% 
  filter(.metric =="roc_auc") %>% 
  arrange(desc(.estimate)) %>% 
  datatable() %>% 
  formatRound(columns = ".estimate",
              digits = 3)
```

Next, I for the model and evaulated its performance on the test data.

```{r}
pca_logistic_tuned_workflow <-
  workflow() %>% 
  add_recipe(rec_pca) %>% 
  add_model(logistic_spec) %>% 
  finalize_workflow(logistic_tune %>%
                      extract_workflow_set_result("pca_logistic") %>%
                      select_best(metric = "roc_auc"))

pca_logistic_fit <-
  pca_logistic_tuned_workflow %>% 
  last_fit(data_split)

pca_logistic_fit %>% collect_metrics()

pred_logistic <-
  pca_logistic_fit %>% 
  collect_predictions() %>% 
  mutate(algo = "PCA Logistic")

pred_logistic %>% 
  roc_curve(outcome, .pred_mutagen) %>% 
  autoplot()

```

Similar to the previous exercise, the Logistic Regression model didn't perform as well on test data. Here's a plot of both roc_auc curves.

```{r}
results <-
  bind_rows(pred_rf, pred_logistic)
```

```{r, eval = TRUE, results = TRUE, preview = TRUE}
results %>% 
  group_by(algo) %>% 
  roc_curve(outcome, .pred_mutagen) %>% 
  autoplot() +
  theme(legend.position = "bottom") +
  theme_bw()
```

