---
title: "Nikkei 225 Option Report 2024-01-31"
date: 2024-01-31

categories:
  - nikkei225
  - options
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    toc_float: true    
    self_contained: false
---

```{r echo = FALSE, include=FALSE}


# Remove all objects in workspace ----
rm(list = ls())
#setwd("C:/Users/User/Dropbox/Papa Work/Data Science/R/SMU/Readings")

# Please import necessary libraries ----
#pacman::p_unload("all")

#### Set up dependencies (Activate necessary packages) ----
pacman::p_load(tidyverse, lubridate,glue, forcats, knitr, rmarkdown,
               scales, gridExtra, ggthemes, ggrepel, cowplot, magick, plotrix, 
               patchwork, hexbin,
               data.table, styler, DT, blogdown,
               quantmod, bizdays, roptions, plotly,
               gtrendsR,rvest, Rcrawler, RCurl, httr2, crul, jsonlite, RSelenium, httr,
               mapdata, maps, ggmap,
               tidytext, wordcloud2, igraph, ggraph, textdata, tm, haven, readxl,
               jtools, huxtable, broom, modelr, skimr, psych, Hmisc, texreg, GGally,
               randomForest, gvlma, ggfortify, sandwich, car, ggstance, broom.mixed,
               interactions, distill
)


theme_set(theme_clean())
sessionInfo()


# Import Data ----
# download daily open interest ----
download_date <-  #"20240118"
  format(today(), "%Y%m%d")
URL <- glue("https://www.jpx.co.jp/english/markets/derivatives/trading-volume/tvdivq00000014nn-att/{download_date}open_interest_e.xlsx")
file_name <- glue("{download_date}open_interest_e.xlsx")
GET(URL, write_disk(file_name, overwrite=TRUE))

data <- read_excel(file_name,
                   sheet = "Attachment1",
                   skip = 4)

names(data)
# download daily option theoretical price ----
URL1 <-glue("https://www.jpx.co.jp/english/markets/derivatives/option-price/data/ose{download_date}tp.zip")
file_name1 <- glue("ose{download_date}tp.zip")
GET(URL1, write_disk(file_name1, overwrite=TRUE))
d1 <- unzip(file_name1)
data1 <- read_csv(d1, col_names = F)

# Extract put data from data ----
data_put <-
  data %>% 
  mutate(contract = as.factor(str_extract(`<Nikkei 225 Options (PUT) >`, pattern = "(?<=P)\\d+")),
         strike = as.numeric(str_extract(`<Nikkei 225 Options (PUT) >`, pattern = "(?<=-)\\d+")),
         type = as.factor(rep("P"))
  ) %>% 
  rename(OI_change = "...4",
         OI_current = "...3",
         volume = "...2") %>%
  select(12:14, 2:4) %>% 
  drop_na() %>% 
  mutate(volume = as.numeric(volume),
         OI_current = as.numeric(OI_current),
         OI_change = as.numeric(OI_change),
         ID = paste0("20", contract, type, strike)
  )
# Extract call data from data ----
data_call <-
  data %>% 
  mutate(contract = as.factor(str_extract(`<Nikkei 225 Options (CALL) >`, pattern = "(?<=C)\\d+")),
         strike = as.numeric(str_extract(`<Nikkei 225 Options (CALL) >`, pattern = "(?<=-)\\d+")),
         type = as.factor(rep("C"))
  ) %>% 
  rename(OI_change = "...10",
         OI_current = "...9",
         volume = "...8") %>% 
  select(12:14, 8:10) %>% 
  drop_na() %>% 
  mutate(volume = as.numeric(volume),
         OI_current = as.numeric(OI_current),
         OI_change = as.numeric(OI_change),
         ID = paste0("20", contract, type, strike)
  )
str(data_call)

# extract Put theoretical price and volatility ----

data_tpv_p <-
  data1 %>% 
  filter(X1 == "NK225E" & X2 == "OOP" & X3 %in% c("202404", "202402", "202403")) %>% 
  mutate(strike = as.numeric(X4), 
         price = as.numeric(X9),
         volatility = as.numeric(X10),
         ID = paste0(X3, "P", strike)
  ) %>% 
  select(19:21)
str(data_tpv_p)
# extract Call theoretical price and volatility ----
data_tpv_c <-
  data1 %>% 
  filter(X1 == "NK225E" & X2 == "OOP" & X3 %in% c("202404", "202402", "202403")) %>% 
  mutate(strike = as.numeric(X4),
         price = as.numeric(X14),
         volatility = as.numeric(X15),
         ID = paste0(X3, "C", strike)
  ) %>% 
  select(19:21)
str(data_tpv_c)
# combine everything into one data set data_options ----

data_tpv <-
  rbind(data_tpv_c, data_tpv_p)

data_call_put <-
  rbind(data_call, data_put)

data_options <-
  full_join(data_tpv, data_call_put, by = "ID") %>% 
  drop_na()

# Create Japan holiday calendar ----
hols <- as.Date(c("2023-11-23", #thanksgiving
                  "2023-12-31", #market holiday
                  "2024-01-01", #market holiday
                  "2024-01-02", #market holiday
                  "2024-01-03", #market holiday
                  "2024-01-08", #coming of age
                  "2024-02-11", #national foundation day
                  "2024-02-12", #national foundation day observed
                  "2024-02-23", #emperors birthday
                  "2024-03-20", #vernal equinox
                  "2024-04-29", #showa day
                  "2024-05-03", #constitution memorial day
                  "2024-05-04", #greenery day
                  "2024-05-05", #childrens day
                  "2024-05-06", #childrens day observed
                  "2024-07-15", #marine day
                  "2024-08-11", #mountain day
                  "2024-08-12", #mountain day observed
                  "2024-09-16", #respect for the aged day
                  "2024-09-22", #autumn equinox
                  "2024-09-23", #autumn equinox observed
                  "2024-10-14", #sports day
                  "2024-11-03", #culture day
                  "2024-11-04", #culture day observed
                  "2024-11-23", #labor thanksgiving day
                  "2024-12-31" #market holiday
)
)
cal <- create.calendar("JPX_JAPAN",
                       holidays = hols,
                       weekdays = c("saturday", "sunday")
)
### Create days to expiration dte for each contract ----
# expressed as T = dte/245 where 245 biz days in 2024

dte_2404 <- bizdays(today(), "2024-04-12", "JPX_JAPAN") / 245
dte_2402 <- bizdays(today(), "2024-02-09", "JPX_JAPAN") / 245
dte_2403 <- bizdays(today(), "2024-03-08", "JPX_JAPAN") / 245

### Staging: Lets extract some relevant information ----
spot_price <- round(as.numeric(data1$X16[1]), digits = 0) # yesterday's closing price
from_strike <- 0.8 * spot_price # choice 20% below spot to view gamma
to_strike <- 1.2 * spot_price # choice 20% above spot to view gamma
spot_price_levels <- seq(from = from_strike, to = to_strike, length.out = 1200)
p_date <- ymd(download_date)

# Staging: getting data for OI plots
d_near_c <- 
  data_options %>%
  filter(type == "C" & contract == "2402")

d_near_p <- 
  data_options %>%
  filter(type == "P" & contract == "2402")

d_near <-
  full_join(d_near_c, d_near_p, by = "strike")

d_far_c <- 
  data_options %>%
  filter(type == "C" & contract != "2402") %>% 
  group_by(strike) %>% 
  summarise(volume = sum(volume),
            OI_current = sum(OI_current),
            OI_change = sum(OI_change))

d_far_p <- 
  data_options %>%
  filter(type == "P" & contract != "2402") %>% 
  group_by(strike) %>% 
  summarise(volume = sum(volume),
            OI_current = sum(OI_current),
            OI_change = sum(OI_change))
d_far <-
  full_join(d_far_c, d_far_p, by = "strike")

### Function for calculating gamma exposure ----
# S: spot price, K: Strike, T: time to expiration,
# r: risk free rate of return (set to zero for now),
# q: dividend yield (set to zero for now), OI: open interest
# 
calcGamma <- function(S, K, vol, t, r, q, OI){
  
  d1 = (log(S/K) + t*(r - q + ((vol^2)/2))) / (vol*sqrt(t))
  gamma = exp(-q*t) * dnorm(d1) / (S * vol * sqrt(t))
  
  ifelse (t == 0 | vol == 0, result <- 0, result <-  OI * 1000 * S^2* 0.01 * gamma )
  
  return(result)
}

# Calculate Spot Gamma Exposure
# where spot is closing price yesterday
str(data_options)
data_options <-
  data_options %>% 
  mutate(dte = as.numeric(ifelse(contract == "2402", dte_2402,
                                 ifelse(contract == "2403", dte_2403, dte_2404)
  )),
  call_gex = ifelse(type == "C",
                    calcGamma(spot_price, strike, volatility, dte, 0, 0, OI_current),
                    0),
  put_gex = ifelse(type == "P",
                   -1*calcGamma(spot_price, strike, volatility, dte, 0, 0, OI_current),
                   0),
  total_gamma = call_gex + put_gex
  
  )
# setting up data for plot----
d<-
  data_options %>% 
  filter(strike>from_strike, strike< to_strike) %>% #plot for +/- 20% of spot
  group_by(contract, strike) %>% 
  summarise(CallGEX = sum(call_gex),
            PutGEX = sum(put_gex),
            TotalGAMMA = sum(total_gamma)
  )
d %>% 
  ggplot(aes(x = strike,
             y = TotalGAMMA)
  ) +
  geom_col()+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")

p_call_put_gamma <-
  d %>% 
  filter(CallGEX !=0 | PutGEX !=0) %>% 
  ggplot()+
  geom_col(aes(x = strike,
               y = CallGEX),
           fill = "blue",
           alpha = 0.8)+
  geom_col(aes(x = strike,
               y = PutGEX),
           fill = "red",
           alpha = 0.8)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(round(from_strike*1.1), round(to_strike*0.9)),
                     breaks = seq(from = round(from_strike*1.1), to = round(to_strike*0.9), by = 250),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "Spot Gamma Exposure ($M Yen/1% move)",
                     labels = label_number(suffix = "M",
                                           scale = 1e-6)
  ) +
  labs(title = "Spot Gamma Exposure ($M Yen/1% move)",
       subtitle = glue("Nikkei 225 {p_date}, Spot Price: {spot_price}"))


### Create plots for daily traded option volume----
# near month

d_near_volume <-
  d_near %>% 
  select(strike, volume.x, volume.y) %>% 
  filter(volume.x >100 | volume.y >100) # arbitary set at 100 for dynamic x axis

p_opt_volume_near <-
  d_near_volume %>% 
  ggplot() + 
  geom_col(aes(x = strike, # plotting for calls (x)
               y = volume.x),
           fill = "blue",
           show.legend = T
  )+
  geom_col(aes(x = strike, #plotting for puts(y)
               y = -volume.y),
           fill = "red",
           show.legend = T)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(round(from_strike*1.1), round(to_strike*0.9)),
                     breaks = seq(from = round(from_strike*1.1), to = round(to_strike*0.9), by = 250),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  scale_y_continuous(name = "volume")+
  labs(title = glue("Nikkei 225 Options Daily Traded Volume - {p_date}"),
       subtitle = glue("Contract: Feb 2024, Spot Price: {spot_price}"),
       caption = "Put volume is plotted on negative y-axis"
  )+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  geom_label(label = "Calls", #label calls
             color = "blue",
             x = spot_price + 1000,
             y = 1000)+
  geom_label (label = "Puts",
              color = "red",
              x = spot_price - 1000,
              y = - 1000)
# far month

d_far_volume <-
  d_far %>% 
  select(strike, volume.x, volume.y) %>% 
  filter(volume.x >50 | volume.y >50)

p_opt_volume_far <-
  d_far_volume %>% 
  ggplot() + 
  geom_col(aes(x = strike, # plotting for calls (x)
               y = volume.x),
           fill = "blue",
           show.legend = T
  )+
  geom_col(aes(x = strike, #plotting for puts(y)
               y = -volume.y),
           fill = "red",
           show.legend = T)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(round(from_strike*1.1), round(to_strike*0.9)),
                     breaks = seq(from = round(from_strike*1.1), to = round(to_strike*0.9), by = 250),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "volume")+
  labs(title = glue("Nikkei 225 Options Daily Traded Volume - {p_date}"),
       subtitle = glue("Non-Front Month Contracts, Spot Price: {spot_price}"),
       caption = "Put volume is plotted on negative y-axis"
  )+
  geom_label(label = "Calls", #label calls
             color = "blue",
             x = spot_price + 1000,
             y = 1000)+
  geom_label (label = "Puts",
              color = "red",
              x = spot_price - 1000,
              y = -1000)


### Create plots for daily change in open interest----
# near month

d_near_oic <-
  d_near %>% 
  select(strike, OI_change.x, OI_change.y) %>% 
  filter(OI_change.x !=0 | OI_change.y !=0) 

p_opt_change_OI_near <-
  d_near_oic %>% 
  ggplot() + 
  geom_col(aes(x = strike, # plotting for calls (x)
               y = OI_change.x),
           fill = "blue",
           show.legend = T
  )+
  geom_col(aes(x = strike, #plotting for puts(y)
               y = OI_change.y),
           fill = "red",
           show.legend = T)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(round(from_strike*1.1), round(to_strike*0.9)),
                     breaks = seq(from = round(from_strike*1.1), to = round(to_strike*0.9), by = 250),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "volume")+
  labs(title = glue("Nikkei 225 Options Daily Change in Open Interest - {p_date}"),
       subtitle = glue("Contract: Feb 2024, Spot Price: {spot_price}")
  )+
  geom_label(label = "Calls", #label calls
             color = "blue",
             x = spot_price +1000,
             y = 200)+
  geom_label (label = "Puts",
              color = "red",
              x = spot_price - 1000,
              y = 200)
# far month

d_far_oic <-
  d_far %>% 
  select(strike, OI_change.x, OI_change.y) %>% 
  filter(OI_change.x !=0 | OI_change.y !=0) 

p_opt_change_OI_far <-
  d_far_oic %>% 
  ggplot() + 
  geom_col(aes(x = strike, # plotting for calls (x)
               y = OI_change.x),
           fill = "blue",
           show.legend = T
  )+
  geom_col(aes(x = strike, #plotting for puts(y)
               y = OI_change.y),
           fill = "red",
           show.legend = T)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(round(from_strike), round(to_strike)),
                     breaks = seq(from = round(from_strike), to = round(to_strike), by = 250),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "volume")+
  labs(title = glue("Nikkei 225 Options Daily Change in Open Interest - {p_date}"),
       subtitle = glue("Non-Front Month Contracts, Spot Price: {spot_price}")
  )+
  geom_label(label = "Calls", #label calls
             color = "blue",
             x = spot_price + 1000,
             y = 400)+
  geom_label (label = "Puts",
              color = "red",
              x = spot_price - 1000,
              y = 400)

### Create plots for daily total open interest----
# near month

d_near_toi <-
  d_near %>% 
  select(strike, OI_current.x, OI_current.y) %>% 
  filter(OI_current.x >500 | OI_current.y >500) # arbitary set at 500 for dynamic x axis


p_opt_OI_current_near <-
  d_near_toi %>% 
  ggplot() + 
  geom_col(aes(x = strike, # plotting for calls (x)
               y = OI_current.x),
           fill = "blue",
           show.legend = T
  )+
  geom_col(aes(x = strike, #plotting for puts(y)
               y = -OI_current.y),
           fill = "red",
           show.legend = T)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(20000, 42000),
                     breaks = seq(from = 20000, to = 42000, by = 500),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "volume")+
  labs(title = glue("Nikkei 225 Options Daily Open Interest - {p_date}"),
       subtitle = glue("Contract: Feb 2024, Spot Price: {spot_price}"),
       caption = "Put volume is plotted on negative y-axis"
  )+
  geom_label(label = "Calls", #label calls
             color = "blue",
             x = spot_price +1000,
             y = 10000)+
  geom_label (label = "Puts",
              color = "red",
              x = spot_price - 1000,
              y = -10000)
# far month

d_far_toi <-
  d_far %>% 
  select(strike, OI_current.x, OI_current.y) %>% 
  filter(OI_current.x >500 | OI_current.y >500) # arbitary set at 500 for dynamic x axis


p_opt_OI_current_far <-
  d_far_toi %>% 
  ggplot() + 
  geom_col(aes(x = strike, # plotting for calls (x)
               y = OI_current.x),
           fill = "blue",
           show.legend = T
  )+
  geom_col(aes(x = strike, #plotting for puts(y)
               y = -OI_current.y),
           fill = "red",
           show.legend = T)+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  scale_x_continuous(name = "strike",
                     limits = c(20000, 42000),
                     breaks = seq(from = 20000, to = 42000, by = 500),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3))+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "volume")+
  labs(title = glue("Nikkei 225 Options Daily Open Interest - {p_date}"),
       subtitle = glue("Non-Front Month Contracts, Spot Price: {spot_price}"),
       caption = "Put volume is plotted on negative y-axis"
  )+
  geom_label(label = "Calls", #label calls
             color = "blue",
             x = spot_price + 1000,
             y = 4000)+
  geom_label (label = "Puts",
              color = "red",
              x = spot_price - 1000,
              y = -4000)


### Here comes the challeging part.
### Calculating Gamma exposure profile ----
# staging
total_gamma <- c()
total_gamma_ex_front <- c()

# For each spot level, calc gamma exposure at that point
for (level in spot_price_levels) {
  
  option_call_gammaEx <-
    data_options %>%
    filter(type == "C") %>%
    mutate(callGammaEx = calcGamma(level, strike, volatility, dte, 0, 0, OI_current)) %>% 
    select(14) %>% 
    sum()
  
  option_put_gammaEx <-
    data_options %>%
    filter(type == "P") %>%
    mutate(putGammaEx = calcGamma(level, strike, volatility, dte, 0, 0, OI_current)) %>% 
    select(14) %>% 
    sum()
  
  total_gamma <- append(total_gamma,option_call_gammaEx-option_put_gammaEx)
  #print(totalGamma)
  
  # calculate gamma ex front month NEED TO UPDATE MONTHLY
  
  option_call_gammaEx_ex_front <-
    data_options %>%
    filter(type == "C" & contract != "2403") %>%
    mutate(callGammaEx = calcGamma(level, strike, volatility, dte, 0, 0, OI_current)) %>% 
    select(14) %>% 
    sum()
  
  option_put_gammaEx_ex_front <-
    data_options %>%
    filter(type == "P" & contract != "2403") %>%
    mutate(putGammaEx = calcGamma(level, strike, volatility, dte, 0, 0, OI_current)) %>% 
    select(14) %>% 
    sum()
  
  total_gamma_ex_front <- append(total_gamma_ex_front,option_call_gammaEx_ex_front-option_put_gammaEx_ex_front)
  #print(total_gamma_ex_front)
  
}

df_gamma <- data.frame(spot_price_levels, total_gamma, total_gamma_ex_front)

### Find where gamma flip occurs ie: when gamma goes from positive to negative ----
# Find Gamma Flip Point for total gamma
zero_cross <- which(diff(sign(df_gamma$total_gamma)) != 0)

neg_gamma <- df_gamma$total_gamma[zero_cross]
pos_gamma <- df_gamma$total_gamma[zero_cross+1]
neg_strike <- df_gamma$spot_price_levels[zero_cross]
pos_strike <- df_gamma$spot_price_levels[zero_cross+1]

zero_gamma <- pos_strike - ((pos_strike - neg_strike) * pos_gamma/(pos_gamma-neg_gamma))

# Find gamma flip point for total gamma ex front month

zero_cross_back <-which(diff(sign(df_gamma$total_gamma_ex_front)) !=0)

neg_gamma_back <- df_gamma$total_gamma_ex_front[zero_cross_back]
pos_gamma_back <- df_gamma$total_gamma_ex_front[zero_cross_back+1]
neg_strike_back <- df_gamma$spot_price_levels[zero_cross_back]
pos_strike_back <- df_gamma$spot_price_levels[zero_cross_back+1]

zero_gamma_back <- pos_strike_back - ((pos_strike_back - neg_strike_back) * pos_gamma_back/(pos_gamma_back-neg_gamma_back))

# plot for total gamma exposure (all months)
p_total_gamma<-
  df_gamma %>% 
  ggplot(aes(x = spot_price_levels)
  )+
  geom_line(aes(y = total_gamma),
            color = "blue"
  )+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  geom_text(aes(x = spot_price,
                y = 0),
            label = round(spot_price, digits = 0),
            color = "black",
            size = 3,
            nudge_x = 400,
            nudge_y = 5e10)+
  geom_vline(xintercept = zero_gamma,
             linetype = "dotdash",
             color = "red")+
  geom_point(aes(x = max(zero_gamma),
                 y = 0),
             color = "red"
  )+
  geom_text(aes(x = max(zero_gamma),
                y = 0),
            label = round(max(zero_gamma), digits = 0),
            size = 3,
            nudge_x = -400,
            nudge_y = 5e10
  )+
  scale_x_continuous(name = "Price",
                     limits = c(spot_price-6000, spot_price+5000),
                     breaks = c(spot_price-6000+500*1:20),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3)
  )+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "Gamma Exposure ($M Yen/1% move)",
                     labels = label_number(suffix = "M",
                                           scale = 1e-6)
  ) +
  geom_hline(yintercept = 0)+
  ggplot2::annotate("rect", xmin = max(zero_gamma), xmax = Inf,
                    ymin = min(total_gamma),ymax = max(total_gamma), 
                    alpha = 0.1, fill = "green") +
  ggplot2::annotate("rect", xmin = -Inf, xmax = max(zero_gamma),
                    ymin = min(total_gamma),ymax = max(total_gamma), 
                    alpha = 0.1, fill = "red")+
  labs(title = glue("Gamma Exposure Profile"),
       subtitle = glue("Nikkei 225 {p_date}, Spot Price: {spot_price}"))


# Plot for total gamma back month


p_total_gamma_back<-
  df_gamma %>% 
  ggplot(aes(x = spot_price_levels)
  )+
  geom_line(aes(y = total_gamma_ex_front),
            color = "blue"
  )+
  geom_vline(xintercept = spot_price,
             linetype = "longdash",
             color = "darkgreen")+
  geom_text(aes(x = spot_price,
                y = 0),
            label = round(spot_price, digits = 0),
            color = "black",
            size = 2,
            nudge_x = 400,
            nudge_y = 5e10)+
  geom_vline(xintercept = zero_gamma_back,
             linetype = "dotdash",
             color = "red")+
  geom_point(aes(x = max(zero_gamma_back),
                 y = 0),
             color = "red"
  )+
  geom_text(aes(x = max(zero_gamma_back),
                y = 0),
            label = round(max(zero_gamma_back), digits = 0),
            size = 2,
            nudge_x = -400,
            nudge_y = 5e10
  )+
  scale_x_continuous(name = "Price",
                     limits = c(from_strike, to_strike),
                     breaks = seq(from = from_strike, to = to_strike, by = 1000),
                     labels = label_number(suffix = "K",
                                           scale = 1e-3)
  )+
  theme(legend.text = element_text(size = rel(0.6)),
        legend.title = element_text(size = rel(0.6)),
        plot.title = element_text(size = rel(0.9)),
        plot.subtitle = element_text(size = rel(0.8)),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(angle = 45, size = rel(0.5)))+
  scale_y_continuous(name = "Gamma Exposure ($M Yen/1% move)",
                     labels = label_number(suffix = "M",
                                           scale = 1e-6)
  ) +
  geom_hline(yintercept = 0)+
  ggplot2::annotate("rect", xmin = max(zero_gamma_back), xmax = Inf,
                    ymin = min(total_gamma_ex_front),ymax = max(total_gamma_ex_front), 
                    alpha = 0.1, fill = "green") +
  ggplot2::annotate("rect", xmin = -Inf, xmax = max(zero_gamma_back),
                    ymin = min(total_gamma_ex_front),ymax = max(total_gamma_ex_front), 
                    alpha = 0.1, fill = "red")+
  labs(title = glue("Gamma Exposure Profile - Excluding Front Month Options"),
       subtitle = glue("Nikkei 225 {p_date}, Spot Price: {spot_price}"))




#ggplotly(p_total_gamma)

```

### Traded Volume:

Date: `r {p_date}` Contract: Feb 2024\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_opt_volume_near
```

Contract: March & April 2024\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_opt_volume_far
```

### Open Interest:

Date: `r {p_date}`Contract: Feb 2024\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_opt_OI_current_near
```

Contract: March & April 2024\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_opt_OI_current_far
```

### Change in Open Interest:

Date: `r {p_date}` Contract: Feb 2024\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_opt_change_OI_near
```

Contract: March & April 2024\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_opt_change_OI_far
```

### Spot Gamma Exposure:

Date: `r {p_date}`\
Nikkei 225 Closing Price: `r {spot_price}`\

```{r, layout="l-page", warning=FALSE, echo = FALSE}
p_call_put_gamma
```

### Gamma Exposure Profile (All Months)

Date: `r {p_date}`\
Nikkei 225 Closing Price: `r {spot_price}`\

```{r, layout="l-page", warning=FALSE, preview = TRUE, echo = FALSE}
p_total_gamma
```

### Gamma Exposure Profile (Excluding Front Month)

Date: `r {p_date}`\
Nikkei 225 Closing Price: `r {spot_price}`\

```{r, layout="l-page", warning=FALSE, preview = TRUE, echo=FALSE}
p_total_gamma_back
```
```{r, layout="l-page", warning=FALSE, preview = TRUE, echo =FALSE}

f <-
  ggplotly(p_total_gamma)
f<-
  f %>% 
  layout(autosize = F,
         widths = 900,
         height = 600)
f
```
