---
title: "k-means clustering using tidyclust and TidyTuesday employment status dataset"
description: |
  A short description of the post.
date: 2024-03-20
categories:
  - unsupervised learning
  - clustering
  - tidyclust
  - k-means
draft: TRUE
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    toc_float: true    
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{css, echo = FALSE}
d-byline {
  display: none;
  margin:0;
}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(results = FALSE, echo = TRUE, eval = FALSE, fig.width = 6, layout="l-body-outset")
```
### Introduction

In module 4 of my class, [Predictive Analytics and Machine Learning Module 4: K-Means and Hierarchical Clustering](https://academy.smu.edu.sg/predictive-analytics-and-machine-learning-module-4-k-means-and-hierarchical-clustering-synchronous) we will be learning about clustering.

I wanted to get a head-start on the material, and as usual, the first place I looked was to my "R-companion's" [Julia Silge's](https://juliasilge.com/blog/kmeans-employment/) blog. Julia's video was published in 2021, likely before Emil Hvitfeldt launched [tidyclust](https://tidyclust.tidymodels.org/index.html).

I shall attempt to walk thru the problem set introduced by Julia, but using `tidyclust`. To recap, the modeling goal is to use k-means clustering to explore employment by race and gender.

### Set dependencies and import necessary packages

```{r}
pacman::p_load(tidyverse, tidymodels, skimr, tidyclust)

```


### Importing the data

Let's begin by importing the dataset from github. I will also save a copy of the data on my computer.

```{r}
df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/employed.csv")
#write_csv(df, "employed.csv")
```

Let's examine the data using `skim`.

```{r}
skim(df)
```











